---
title: "Initial Exploration of Assistance tracking in RAIS - Iraq-Jordan-Lebanon"
author: "Prepared by UNHCR DOiA, MENA Protection Services"
date: "Amman, `r Sys.Date()`"
output: 
  word_document: 
    fig_caption: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
setwd("D:/R-project/assistance-analysis")

source("code/-packages.R")

assistance.case <- fread( file = "data/assistancecase.csv",na="")


```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This reports presents the analysis of assistance recorded in the RAIS - (Refugee Assistance Information System) fro Jordan, Lebanon & Iraq.

It represents a total of `r nrow(data)` :   `r nrow(data[data$instance=="Iraq", ])`  for Iraq, `r nrow(data[data$instance=="Lebanon", ])`  for Lebanon and  `r nrow(data[data$instance=="Jordan", ])`  for Jordan.



Assistance Type were recategorised in order to come with meaningfull categories for analysis.

# Assistance record per country over years


# Assistancetype, country and months

# Details of cash assistance

# Details of NFI assistance

# 

# Assistance Type from & through - by year & country
The following sankey diagram provided an overview of the origin and delivery stream of 


# Assistance occurence by case
The following visualisation provides an overview of the most associated assistance type



